Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm

### 2019-10-05 paper/virtual_cytoplasm_v2.tex
Updates to paper adding many questions.

\documentclass[journal=jacsat,manuscript=article]{achemso}
% \linespread{1.5}
\usepackage{amssymb,amsmath,graphics,epsfig}
% \usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

\begin{abstract}
Biomolecular crowding affects the biophysical and biochemical behavior of macromolecules when compared to the dilute environment present in experiments made with isolated proteins. Computational modeling and simulation are essential tools in the investigation of how crowding affects the structural dynamics and biological activities of macromolecules. As computational power increased, modeling and simulating large scale all-atom explicit solvent models of the prokaryote cytoplasm become possible. In this work, we build an atomistic model of the cytoplasm of \textit{Escherichia coli} composed of 1.5 million atoms and submit it to a total of \SI{3}{\micro\second} of molecular dynamics simulations, which is unprecedented for a system of this scale. Our model reproduces experimental data about the diffusion coefficient and stability of macromolecules under crowded conditions. We provide a Python script and a set of files that enables any researcher to build their own \textit{E. coli} cytoplasm models to address questions related to crowding.
%For Nature, the abstract is really an introductory paragraph set
%in bold type.  This paragraph must be ``fully referenced'' and
%less than 180 words for Letters.  This is the thing that is
%supposed to be aimed at people from other disciplines and is
%arguably the most important part to getting your paper past the
%editors.  End this paragraph with a sentence like ``Here we
%show...'' or something similar.

\end{abstract}
\section*{Introduction}

@@

The intracellular environment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}. Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}. This effect, called macromolecular crowding, directly influences the stability of macromolecules through an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}. Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013}.

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it was estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radii to represent different types of molecules~\cite{Ando2010}. These models usually use an implicit representation of the solvent~\cite{Mcguffee2010}. More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed of several copies of one or a few macromolecules~\cite{Wang2017c}. However, these models do not reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms~\cite{Feig2015}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations~\cite{Yu2016a}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize dynamical effects that can take place in such a large scale.

In this work, we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system. We discuss the difficulties that arise when building such system and provide a set of Python scripts and files that anyone can use to build its own crowded systems with custom parameters and elements. Additionally, we performed long molecular dynamics simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of total simulation time. This work is the first to explore the structural dynamics of an all-atom explicit solvent cytoplasm model with microsecond scale molecular dynamics simulations.

@@

\subsection{Initial structures}
The proteins and tRNA were downloaded from Protein Data Bank (PDB). We looked for the proteins structures that were either from {\em E. coli} or its closest homologues (Table \ref{tbl:protein_fraction}). In case of PDB IDs 1U22 (MetE protein) and 2EIP (Ppa protein), we used a loop-closure modeling tool based on the Random Coordinate Descent (RCD) method~\cite{Chys2013} to model missing residues. The four metabolites were parametrized using the General Amber Force Field (GAFF)~\cite{sprenger2015general} and Antechamber~\cite{Wang2005b} using RESP charges~\cite{Bayly1993a} that were derived from electron density calculated at the HF/6-31G* theory level with Gaussian16.

\subsection{General simulation setup}
The cytoplasm model was built at 30\% biomolecular mass fraction at the physiological salt concentration of 0.15 mol/L KCl. All simulations were performed with the GROMACS 2018 package~\cite{Abraham2015a} with Amber99SB-ws~\cite{Best2014a}, a modified version of the Amber99SB force field, in combination with the TIP4P/2005 water model~\cite{Abascal2005b}. All interactions were calculated explicitly inside a 1 nm radius and long range electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a}. Corrections due to long range effects of dispersion interactions were made to both energy and pressure. All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs. Temperature was kept at 310 K by the v-rescale thermostat~\cite{Bussi2007a} with a coupling constant of 0.5 ps. Pressure was kept at 1 bar by the Berendsen barostat~\cite{Berendsen1984a} with the coupling constant of 10.0 ps during thermalization steps and by the Parrinello-Rahman barostat~\cite{Parrinello1981a} with the same coupling constant during production steps.

All systems were initially submitted to energy minimization with the steepest descent algorithm until convergence to simple precision. For the cytoplasm model, a short shrinking step of 500 ps was performed in order to allow the box volume to adjust to its optimal value. For all systems, initial velocities were sampled from a Maxwell-Boltzmann distribution at 310 K and we ran a 1 ns NVT thermalization step followed by 10 ns of NPT equilibration. Production runs for the dilute systems, in which each crowder is simulated alone in a water box, had the length of 200 ns. The cytoplasm model was run for 1 \SI{}{\micro\second}. All systems were simulated in three replicas. While the replicas for the diluted systems were generated by different initial velocities, each replica of the cytoplasm model is a completely different box in which the components of the model are in different positions and orientations.

\subsection{Analyses}
For statistics, the average and standard error of the three trajectories are reported. For values that are specific to each chain, we used the average over the individual chains in each trajectory of the replicas. Additionally, since there is more than one copy of some crowders in the cytoplasm models, average values and standard errors are calculated considering each copy and each replica independently. Similarly to the dilute system, separate chains of each copy of the crowders were considered as independent samples when running analyses that focus on one single chain. The first half of all simulations was discarded as equilibration time and analyses used only the second half.

Analyses were done by GROMACS tools after removing artifacts generated by periodic boundary conditions from the trajectories. The Mean Square Displacement (MSD) analysis was performed to calculate the translational diffusion coefficient, which were extracted by a linear fit to MSD by averaging blocks with a length of $10 \,ns$~\cite{Allen1987a}. In principle, diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible. Root Mean Square Deviation (RMSD) and Root Mean Square Fluctuation (RMSF) calculations considered only C$\upalpha$ atoms and the reference frame for RMSF was the midpoint of the time interval used for analysis in each trajectory. The chain interface area of a given oligomer in a given frame was calculated as the difference between the solvent exposed surface area (SASA) of the oligomer in that frame and the sum of the SASAs of each of its chains independently in that same frame. Other analyses, such as box volume and Radial Distribution Function (RDF), followed standard procedures.

@@

\subsection{Cytoplasm model}
In this section we describe the rationale behind the composition of our model, which has five fractions: protein, RNA, metabolites, water and ions. We highlight that we didn't add lipids and DNA to our model because we are considering only elements that are free to diffuse through the cytosol. We gathered data from several sources in order to build a computational model that is representative of the cytoplasm of {\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.

\subsubsection{Protein fraction}
We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{Link1997}. The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E. coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}). Their crystallographic structures were used to insert them in the cytoplasm model. The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1. Then, this number was divided by the oligomeric state of each protein. This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure (Table~\ref{tbl:protein_fraction}).

\begin{table}[H]
\centering
\begin{tabular}{ l c c c}
\hline
\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} & \multirow{2}{*}{Oligomeric State}\\
& Absolute & Cumulative \\ 
\hline
TufA & 19.7 & 19.7 & 2-mer \\
MetE & 11.6 & 31.4 & 1-mer \\
IcdA & 4.7  & 36.1 & 2-mer \\
AhpC & 4.1  & 40.2 & 10-mer \\
CspC & 4.0  & 44.2 & 1-mer \\
Ppa  & 2.9  & 47.0 & 6-mer \\
GapA & 2.1  & 49.1 & 4-mer \\
Eno  & 1.9  & 51.1 & 2-mer \\
\hline
\end{tabular}
\caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E. coli} K12 and oligomeric state of each protein that was selected to compose our cytoplasm model.}
\label{tbl:protein_fraction}
\end{table}

\subsubsection{RNA fraction}
Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs~\cite{phillips2012}. Thus, we chose to model the RNA presence in the cytoplasm with tRNA molecules. Specifically, we considered the tRNA(Phe) molecule as a representative of tRNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively~\cite{phillips2012}. That is, the total RNA weight corresponds to 5\% of the total protein weight. This protein/RNA weight ratio was used to calculate the correct number of tRNA(Phe) molecules that were added to the cytoplasm model \tabref{tbl:soup_components}.

\subsubsection{Metabolites fraction}
We considered the most abundant molecules of each metabolite class as representatives, i.e. Glutamate (GLT) for amino acids, ATP for nucleotides, Fructose-2,6-Biphosphate (FBP) for central carbon intermediates and Glutathione (GSH) redox cofactors~\cite{Bennett2009}. The total number of molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E. coli} is about 42.86 times higher than the number of proteins~\cite{Bennett2009}. The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in {\em E. coli}~\cite{Bennett2009}.

\subsubsection{Water fraction}
The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model. This number ranges from 300 to 400 g/L in biological systems such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, which corresponds to biomolecular concentrations as high as 40\%~\cite{Ellis2003a}. In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.

@@

\begin{table}
\centering
\begin{tabular}{lcc}
\hline
Class & Name (PDB ID) & Number\\
\hline
Protein & TufA (1DG1~\cite{Abel1996}) & 6\\
  & MetE (1U22~\cite{Ferrer2004}) & 7\\
  & IcdA (1P8F~\cite{Mesecar2000}) & 2\\
  & AhpC (1YEP~\cite{Parsonage2005}) & 1\\
  & CspC (1MJC~\cite{Schindelin1994}) & 3\\
  & Ppa (2EIP~\cite{Kankare1996}) & 1\\
  & GapA (1S7C~\cite{ShinXXX}) & 1\\
  & Eno (1E9I~\cite{Kuhnel2001}) & 1\\
\hline
RNA & tRNA$^{\text{Phe}}$ (4YCO~\cite{Byrne2015}) & 5\\
\hline
Metabolite & GLT & 1436\\
  & ATP & 144\\
  & FBP & 225\\
  & GSH & 255\\
\hline
Solvent & Water & 306221\\
\hline
Inorganic Ion & K$^{+}$ & 4602\\
  & Mg$^{2+}$ & 400\\
  & Cl$^{-}$ & 1320\\
\hline
\end{tabular}
\caption{Number of copies for each component of the cytoplasm model built at the biomolecular fraction of 30\%.}
\label{tbl:soup_components}
\end{table}

@@

In order to do this, the droplets around each component are also constructed iteratively. The number of water molecules that we must place in the simulation box is known from the total biomolecular mass of the system and the desired biomolecular fraction, which is a parameter of the model. However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules \textit{a priori}. Since $l$ depends on a series of factors such as the shape, size and abundance of each element, we define it iteratively. We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets. If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}. If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increase $l$ with the new smaller $dl$. We can carry this process until an arbitrary precision cutoff, such as 5\%, for the number of waters in the system is satisfied.

After all droplets are successfully inserted in a simulation box by the iterative process described above, we proceed to add ions to neutralize the net charge of the system and to reach the desired ionic strength. Then, we perform energy minimization and a short simulation step of 500ps in which all molecules of the box are free to move. In this step the box shrinks to its optimum size. In our case, the simulation box shrank from the initial box size of 35 nm to 22.9 nm, which correspond to a volume change from 42875 nm\textsuperscript{3} to 12009 nm\textsuperscript{3}. From this point, the system is ready to be submitted to the default simulation steps such as thermalization and production run (please check the materials section for details about the parameters we used).

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking.pdf} 
\caption{A) The initial volume of the simulation box, $L=35.0$ nm, and after the shrinking step, $L=22.9$ nm. B) The evolution of the volume of the simulation box during the shrinking simulations. Each replica is shown in different colors. }
\end{figure}

@@

Under crowded conditions, molecules are constrained to a smaller volume by the other components of the crowded environment. The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Fig.~\ref{fig:translational_diffusion}A). The ratio of these quantities is independent of the protein size and is always close to 0.13 (Fig.~\ref{fig:translational_diffusion}B). However, the effect of crowding on the translational diffusion of tRNA is two times larger. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa).

@@

\subsubsection{tRNA is aggregating}
After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, we will omit the tRNA molecules in the analyses about the structural integrity of the crowders. In the next sections we will further investigate such aggregation and we will show a way to prevent it.

>>>>

\documentclass[journal=jcisd8,manuscript=article]{achemso}
% \linespread{1.5}
\usepackage{amssymb,amsmath,graphics,epsfig}
% \usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

\newpage
\begin{abstract}
Biomolecular crowding affects the biophysical and biochemical behavior
of macromolecules when compared to the dilute environment present in
experiments made with isolated proteins. Computational modeling and
simulation are useful tools to study how crowding affects the
structural dynamics and biological properties of macromolecules. As
computational power increased, modeling and simulating large scale
all-atom explicit solvent models of the prokaryote cytoplasm become
possible. In this work, we build an atomistic model of the cytoplasm
of \textit{Escherichia coli} composed of 1.5 million atoms and submit
it to a total of \SI{3}{\micro\second} of molecular dynamics
simulations. The simulation model is consistent with experimental data
about the diffusion coefficient and stability of macromolecules under
crowded conditions. In order to stimulate further work we provide a
Python script and a set of files that enables any researcher to build
their own \textit{E. coli} cytoplasm models to address questions
related to crowding.
\end{abstract}

\newpage
\section*{Introduction}

@@

The intracellular environment is markedly different from the
\textit{in vitro} model systems from which most of our knowledge about
the biochemical and biophysical behavior of macromolecules has been
derived~\cite{Feig2017a}. The \textit{in vivo} microenvironment is
heterogeneous with high concentrations of many different compounds,
leading to spatial constrains that directly affects biological
activities~\citep{ostrowska2019}. This effect, called macromolecular
crowding, influences the stability of {\bf SOME/MANY} macromolecules
because the equilibrium is shifted towards a stable native structure
as a result of the entropic component given by the excluded volume
effect~\cite{cheung2005}. At the same time, an effective
destabilization due to the enthalpic component of the Gibbs energy has
been observed, due to replacing interactions with water by weaker
interactions with the
crowders~\cite{Feig2011,miklos2011,Wang2012b}. Additionally, crowding
affects the structure and dynamics of water around proteins, which has
a significant effect on their biological
activities~\cite{Harada2012a,king2013}.

While the structure and dynamics of biomolecules is well characterized
\textit{in vitro}, the understanding of the influences of crowded
environments \textit{in vivo} are still evolving.  Techniques such as
nuclear magnetic resonance~\cite{reckel2007,pielak2008} and
fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are
among the most relevant to the field due to their ability to probe
cells in a non destructive way. Another essential class of methods
come from computational modeling and simulation. By extrapolating the
advances in computational power in the last decades to the future, it
was estimated that the atomistic simulation of an entire
\textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be
possible in 15 years from now~\cite{vanGunsteren2006a}. {\bf THAT WAS
  13 YEARS AGO. WHAT IS THE STATE OF THE ART?}

As the available computational power increases, models become
increasingly complex and are better representations of reality. The
most simple class of computational models apply Brownian Dynamics
calculations to probe macromolecules immersed in a solution of
crowders that are treated as hard spheres with varying radii to
represent different types of molecules~\cite{Ando2010}. These models
usually use an implicit representation of the
solvent~\cite{Mcguffee2010}. More recently, all-atom explicit solvent
molecular dynamics simulations have been employed on crowded systems
composed of several copies of one or a few
macromolecules~\cite{Wang2017c}. However, these models do not reflect
the true heterogeneity of biologically relevant crowded environments
such as the cytoplasm. This was addressed by the development of a
detailed model for the cytoplasm of the bacterium \textit{Mycoplasma
  genitalium} containing 103 million atoms~\cite{Feig2015}. This
model, as well as a smaller version with 13 million atoms, was
submitted to molecular dynamics simulations~\cite{Yu2016a}. Due to the
size of these systems, the simulation times were very short,
\SI{20}{\nano\second} for the complete model and
\SI{140}{\nano\second} for a smaller version, precluding an accurate
characterization of dynamical effects that can take place in such a
large scale. It was noted already some years ago that even
{microsecond} simulations may be too short to probe the dynamics of
large complexes like viruses~\cite{Larsson2012a} since dynamics is
slower in larger systems in general.

In this work, we report an all-atom model of part of the cytoplasm of
\textit{Escherichia coli} was built to reflect the real biological
system, in particular as regards a realistic composition of this
``soup''. We discuss the difficulties that arise when building such a
system and provide a set of Python scripts and files that anyone can
use to build its own crowded systems with custom parameters and
compounds. Additionally, we performed long molecular dynamics
simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of
total simulation time. This work is the first to explore the
structural dynamics of an all-atom explicit solvent cytoplasm model
with microsecond scale molecular dynamics simulations. {\bf IS THAT
  REALLY TRUE? A REFEREE MAY GET ANGRY WHEN READING THIS.}

@@

\subsection{Initial structures}
The protein and tRNA structures were downloaded from the Protein Data
Bank (PDB)~\cite{pdb}. Proteins structures that were either from {\em
  E. coli} or its closest homologues (Table
\ref{tbl:protein_fraction}) were selected for this work. In case of
PDB IDs 1U22 (MetE protein) and 2EIP (Ppa protein), we used a
loop-closure modeling tool based on the Random Coordinate Descent
(RCD) method~\cite{Chys2013} to model missing residues. The four
metabolites were parametrized using the General Amber Force Field
(GAFF)~\cite{Wang2004a} and Antechamber~\cite{Wang2005b} using RESP
charges~\cite{Bayly1993a} that were derived from electron density
calculated at the HF/6-31G* theory level with Gaussian16.  Properties
of compounds modeled using the GAFF have been evaluated in a number of
papers~\cite{Caleman2012a,sprenger2015general,Fischer2015a,JZhang2015a,JZhang2015,Spoel2018a}
and strengths and weaknesses are well understood. Here, GAFF is used
for compatibility with the Amber99SB-ws force field~\cite{Best2014a}
used for the biomolecules. {\bf DESCRIBE WHICH MODELS WERE USED FOR
  THE IONS.}

\subsection{General simulation setup}
The cytoplasm model was built at a 30\% biomolecular mass fraction and
a physiological salt concentration of 0.15 mol/L KCl.  {\bf DESCRIBE
  BRIEFLY HOW COMPOUNDS WERE PLACED IN THE BOX AND THAT THE VOLUME WAS
  WAY LARGER. DO ALSO WRITE WHETHER MOLECULES WERE ROTATED randomly
  PRIOR TO PLACEMENT IN THE BOX. ALTERNATIVELY REFER TO RESULTS.}  All
simulations were performed with the GROMACS 2018
package~\cite{Abraham2015a,Pronk2013a} using the Amber99SB-ws force
field~\cite{Best2014a}, a modified version of the Amber99SB force
field, in combination with the TIP4P/2005 water
model~\cite{Abascal2005b}. All interactions were calculated explicitly
inside a 1 nm radius and long range electrostatic interactions were
treated using the particle mesh Ewald
algorithm~\cite{Essmann1995a}. Corrections due to long range effects
of dispersion interactions were made to both energy and
pressure~\cite{Allen1987a}. All chemical bonds were constrained at
their equilibrium length using the parallel LINCS
algorithm~\cite{Hess1997a,Hess2008b} allowing an integration time step
of 2 fs. Temperature was kept at 310 K by the v-rescale
thermostat~\cite{Bussi2007a} with a coupling constant of 0.5
ps. Pressure was kept at 1 bar using the Berendsen
barostat~\cite{Berendsen1984a} with the coupling constant of 10.0 ps
during thermalization steps and by the Parrinello-Rahman
barostat~\cite{Parrinello1981a} with the same coupling constant during
production steps.

All systems were initially submitted to energy minimization with the
steepest descent algorithm until convergence to machine (single)
precision. For the cytoplasm model, a short shrinking step of 500 ps
was performed in order to allow the box volume to adjust to its
optimal value. For all systems, initial velocities were sampled from a
Maxwell-Boltzmann distribution at 310 K and we ran a 1 ns NVT
thermalization step followed by 10 ns of NPT equilibration. Production
runs for the dilute systems, in which each crowder is simulated alone
in a water box, had a- length of 200 ns. The cytoplasm model was run
for 1 \SI{}{\micro\second}. All systems were simulated in three
replicas. While the replicas for the diluted systems were generated by
different initial velocities, each replica of the cytoplasm model is a
completely different box in which the components of the model are in
different positions and orientations.

\subsection{Analyses}
For statistics, the average and standard error {\bf standard error or
  deviation?} of the three trajectories are reported. For values that
are specific to each chain, we used the average over the individual
chains in each trajectory of the replicas. Additionally, since there
is more than one copy of some crowders in the cytoplasm models,
average values and standard errors are calculated {\bf DO YOU MEAN THE
  AVERAGE OVER ALL COPIES? considering each copy and each replica}
independently. As in the dilute system, separate chains of each copy
of the crowders were considered as independent samples when running
analyses that focus on one single chain. The first half of all
simulations was discarded as equilibration time and analyses used only
the second half.

Analyses were done using GROMACS tools after removing artifacts
generated by periodic boundary conditions from the trajectories. A
Mean Square Displacement (MSD) analysis was performed to calculate the
translational diffusion coefficient, which were extracted by a linear
fit to MSD by averaging blocks with a length of {\bf 10 ns even for
  proteins?}  $10\,ns$~\cite{Allen1987a}. In principle, diffusion
coefficient needs to be corrected for finite size
effects~\cite{Yeh2004} but due to relatively large simulation boxes
this correction is negligible {\bf NOT FOR PROTEINS THAT DIFFUSE
  SLOWLY!}. Root Mean Square Deviation (RMSD) and Root Mean Square
Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only
and the reference frame for RMSF was the midpoint of the time interval
{\bf ARE YOU SURE? THE DEFAULT IS TO TAKE THE AVERAGE STRUCTURE. OR DO
  YOU MEAN THE STRUCTURE FOR FITTING?} used for analysis in each
trajectory. The chain interface area of a given oligomer in a given
frame was calculated as the difference between the solvent exposed
surface area (SASA) of the oligomer in that frame and the sum of the
SASAs of each of its chains independently in that same frame. Other
analyses, such as box volume and Radial Distribution Function (RDF),
followed standard procedures.

@@

\subsection{Cytoplasm model}
In this section we describe the rationale behind the composition of
our model, which consists of five fractions: protein, RNA, metabolites, water
and ions. It should be noted that we did not add lipids and DNA to our
model because we consider only elements that are free to diffuse
through the cytosol. We gathered data from several sources in order to
build a computational model that is representative of the cytoplasm of
{\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.

\subsubsection{Protein fraction}
Eight different proteins were selected that together account for 50\%
of the non-ribosomal proteins in the cytoplasm of {\em Escherichia
  coli} \cite{Link1997}. The two most abundant proteins, TufA and
MetE, account for about 20\% and 12\% of the total protein abundance
in {\em E. coli}, respectively, while the other six proteins
contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}).
The least abundant protein was present with 1 copy and the amount of
each of the other proteins was taken to be an integer number of
oligomers. This was done because the original oligomeric state for
each protein was kept as reported in their crystallographic structure
(Table~\ref{tbl:protein_fraction}).

\begin{table}[H]
\centering
\caption{Protein, PDB ID, number N of proteins in the simulation. 
Fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em
    E. coli} K12 and oligomeric state of each protein that was
  selected to compose our cytoplasm model. {\bf WRITE OUT WHOLE
    PROTEIN NAMES}}
\label{tbl:protein_fraction}
\begin{tabular}{lcccc}
\hline
Protein & PDB & N & Fraction [\%] & State\\
\hline
TufA & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
MetE & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
IcdA & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
AhpC & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
CspC & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
Ppa  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
GapA & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
Eno  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
\hline
Total &                           & 48& 51.1 & \\
\hline
\end{tabular}
\end{table}

\subsubsection{RNA fraction}
Transporter RNAs (tRNAs) account for 74\% of the dry weight of
non-ribosomal RNAs~\cite{phillips2012}. Thus, we chose to model the
RNA presence in the cytoplasm with tRNA molecules. Specifically, 
 the tRNA(Phe) molecule as a representative of tRNAs due to
the availability of a recent crystal
structure (4YCO~\cite{Byrne2015}). The protein and RNA content of the total
dry weight of \textit{E.coli} is 55\% and 2.9\%,
respectively~\cite{phillips2012}. That is, the total RNA weight
corresponds to 5\% of the total protein weight. This protein/RNA
weight ratio was used to calculate the correct number of tRNA(Phe)
molecules that were added to the cytoplasm model
\tabref{tbl:soup_components}.

\subsubsection{Metabolites fraction}
The most abundant molecules of each metabolite class was used to 
represent the class, i.e. Glutamate (GLT) for amino acids, ATP for
nucleotides, Fructose-2,6-Biphosphate (FBP) for central carbon
intermediates and Glutathione (GSH) redox
cofactors~\cite{Bennett2009}. The total number of molecules was
calculated considering data showing that the number of metabolite
molecules in the cytoplasm of \textit{E. coli} is $\approx$ 43 times
higher than the number of proteins~\cite{Bennett2009}. The copy number
for each molecule was calculated from the ratios between their
experimentally observed concentration in {\em
  E. coli}~\cite{Bennett2009}.

\subsubsection{Water fraction}
The number of water molecules was calculated according to the desired
biomolecular concentration, which is a parameter of the cytoplasm
model. This number ranges from 300 to 400 g/L in biological systems
such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, which {\bf
  WHAT DOES WHICH CORRESPOND TO?}  corresponds to biomolecular
concentrations as high as 40\%~\cite{Ellis2003a}. In our case, we
choose a biomolecular concentration of 30\%, that is, the number of
molecules necessary to reach a ratio of total biomolecular mass to
water mass of 30\% was inserted into the cytoplasm model.

@@

\begin{table}
\centering
\begin{tabular}{lcc}
\hline
Class & Name & Number\\
\hline
RNA & tRNA$^{\text{Phe}}$ & 5\\
\hline
Metabolite & GLT & 1436\\
  & ATP & 144\\
  & FBP & 225\\
  & GSH & 255\\
\hline
Solvent & Water & 306221\\
\hline
Inorganic Ion & K$^{+}$ & 4602\\
  & Mg$^{2+}$ & 400\\
  & Cl$^{-}$ & 1320\\
\hline
\end{tabular}
\caption{Number of copies for non-protein components of the cytoplasm model built at the biomolecular fraction of 30\%.}
\label{tbl:soup_components}
\end{table}

@@

In order to do this, the droplets around each component are also
constructed iteratively. The number of water molecules that we must
place in the simulation box is known from the total biomolecular mass
of the system and the desired biomolecular fraction, which is a
parameter of the model. However, we don't know the thickness of the
water layer around each component, $l$, that accounts for such amount
of water molecules \textit{a priori}. Since $l$ depends on a series of
factors such as the shape, size and abundance of each element, we
define it iteratively. We start by taking a droplet of thickness $l$ =
3 \r{A} around all elements and counting the number of water molecules
that we would add to the box with such droplets. If it is smaller than
the number we need, $l$ is increased by $dl$ = 1 \r{A}. If it is
larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10
times and then increase $l$ with the new smaller $dl$. We can carry
this process until a user-defined precision cutoff, here 5\%, for the
number of waters in the system is satisfied.

After all droplets are successfully inserted in a simulation box by
the iterative process described above, we proceed to add ions to
neutralize the net charge of the system and to reach the desired ionic
strength. Then, we perform energy minimization and a short simulation
 of 500ps in which all molecules of the box are free to move. In
this step the box shrinks to its optimum size. In our case, the
simulation box shrank from the initial box size of 35 nm to 22.9 nm,
which corresponds to a volume change from 42875 nm\textsuperscript{3}
to 12009 nm\textsuperscript{3}. From this point, the system is ready
to be submitted to the default simulation steps such as thermalization
and production run (please check the materials section for details
about the parameters used).

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking.pdf} 
\caption{A) The initial simulation box with $L=35.0$ nm and the box
  after the shrinking step, $L=22.9$ nm. B) The evolution of the
  volume of the simulation box during the shrinking simulations. Each
  replica is shown in different colors. }
\end{figure}

@@

Under crowded conditions, molecules are confined to a smaller
effective volume by the other components of the environment.  The
extent of this effect is evident when we compare the translational
diffusion constant of each component in the cytoplasm model,
$D_{cyto}$, and in a dilute system, $D_{dil}$
(Fig.~\ref{fig:translational_diffusion}A). The ratio of these
quantities is independent of the protein size and is always close to
0.13 (Fig.~\ref{fig:translational_diffusion}B). However, the effect of
crowding on the translational diffusion of tRNA is two times
larger. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement
is especially constrained in the crowded environment
(Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa),
however, see below.

@@

\subsubsection{tRNA is aggregating}
After inspecting the trajectories of the cytoplasm models, we found
that the reason why the translational diffusion constant for tRNA was
reduced to a greater extent than for the other components of the
cytoplasm model is that it was forming aggregates with Mg$^{2+}$, ATP
and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, we will omit the tRNA
molecules in the analyses about the structural integrity of the
crowders. In the next sections we will further investigate such
aggregation and we will show a way to prevent it.
